---
title: "PHP2560 Pset #1 (Week 02 Assignment)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1:

1. Generate 200 random values from the standard exponential distribution and store them in a vector exp.draws.1. Find the mean and standard deviation of exp.draws.1.

```
#Generate 200 random values from the standard exponential dist.
#standard exponential distribution has lambda = 1. 
exp.draws.1 = rexp(200)
mean_1 = mean(exp.draws.1)
sd_1 = sd(exp.draws.1)
```

2. Repeat, but change the rate to 0.2, 5, 7.3 and 10, storing the results in vectors called exp.draws.0.2, exp.draws.5, exp.draws.7.3 and exp.draws.10.

```
#Repeat for rates lambda = 0.2
exp.draws.0.2 = rexp(200, rate = 0.2)
mean_0.2 = mean(exp.draws.0.2)
sd_0.2 = sd(exp.draws.0.2)

#lambda = 5
exp.draws.5 = rexp(200, rate = 5)
mean_5 = mean(exp.draws.5)
sd_5 = sd(exp.draws.5)

#lambda = 7.3 
exp.draws.7.3 = rexp(200, rate = 7.3)
mean_7.3 = mean(exp.draws.7.3)
sd_7.3 = sd(exp.draws.7.3)

#lambda = 10
exp.draws.10 = rexp(200, rate = 10)
mean_10 = mean(exp.draws.10)
sd_10 = sd(exp.draws.10)
```

3. The function plot() is the generic function in R for the visual display of data. hist() is a function that takes in and bins data as a side effect. To use this function, we must first specify what we'd like to plot. a. Use the hist() function to produce a histogram of your standard exponential distribution. b. Use plot() with this vector to display the random values from your standard distribution in order. c. Now, use plot() with two arguments -- any two of your other stored random value vectors -- to create a scatterplot of the two vectors against each other.

```
#Histogram of standard exponential distribution
hist(exp.draws.1)

#Plot values generated by the standard exponential dist. in the order they were generated
plot(exp.draws.1)

#Scatter plot between the standard distribution and the exponential distribution with lambda = 10
plot(exp.draws.1, exp.draws.10)
```
4. We'd now like to compare the properties of each of our vectors. Begin by creating a vector of the means of each of our five distributions in the order we created them and saving this to a variable name of your choice. Using this and other similar vectors, create the following scatterplots and explain in words what is going on: a. The five means versus the five rates used to generate the distribution. b. The standard deviations versus the rates. c. The means versus the standard deviations.

For each plot, explain in words what's going on.

```
#Scatterplots from mean vector of previous distribution
mean_vector = c(mean_1, mean_0.2, mean_5, mean_7.3, mean_10)
sd_vector = c(sd_1, sd_0.2, sd_5, sd_7.3, sd_10)
rate_vector = c(1, 0.2, 5, 7.3, 10)
plot(rate_vector, mean_vector) #part a
plot(rate_vector, sd_vector) #part b
plot(sd_vector, mean_vector) #part c
```

## Part 2:

5. R's capacity for data and computation is large to what was available 10 years ago. a. To show this, generate 1.1 million numbers from the standard exponential distribution and store them in a vector called big.exp.draws.1. Calculate the mean and standard deviation. b. Plot a histogram of big.exp.draws.1. Does it match the function (1-e^{-x})? Should it? c. Find the mean of all of the entries in big.exp.draws.1 which are strictly greater than 1. You may need to first create a new vector to identify which elements satisfy this. d. Create a matrix, big.exp.draws.1.mat, containing the the values in big.exp.draws.1, with 1100 rows and 1000 columns. Use this matrix as the input to the hist() function and save the result to a variable of your choice. What happens to your data? e. Calculate the mean of the 371st column of big.exp.draws.1.mat. f. Now, find the means of all 1000 columns of big.exp.draws.1.mat simultaneously. Plot the histogram of column means. Explain why its shape does not match the histogram in problem 5b).

```
#Part A:
big.exp.draws.1 = rexp(1100000)
big_mean = mean(big.exp.draws.1)
big_sd = sd(big.exp.draws.1)

#Part B:
hist(big.exp.draws.1)
x = seq(0, 15, by = 0.1)
plot(x, 1 - exp(-x))

#Part C:
mean(big.exp.draws.1[which(big.exp.draws.1 > 1)])

#Part D:
big.exp.draws.1.mat = matrix(data = big.exp.draws.1, nrow = 1100, ncol = 1000)
big_hist = hist(big.exp.draws.1.mat)

#Part E: 
col_371_mean = mean(big.exp.draws.1.mat[, 371])

#Part F:
col_means = colMeans(big.exp.draws.1.mat)
hist(col_means)

```

## Including Plots
```{r, echo=FALSE}
#Generate 200 random values from the standard exponential dist.
#standard exponential distribution has lambda = 1. 
exp.draws.1 = rexp(200)
mean_1 = mean(exp.draws.1)
sd_1 = sd(exp.draws.1)
median_1 = median(exp.draws.1)

#Repeat for rates lambda = 0.2
exp.draws.0.2 = rexp(200, rate = 0.2)
mean_0.2 = mean(exp.draws.0.2)
sd_0.2 = sd(exp.draws.0.2)

#lambda = 5
exp.draws.5 = rexp(200, rate = 5)
mean_5 = mean(exp.draws.5)
sd_5 = sd(exp.draws.5)

#lambda = 7.3
exp.draws.7.3 = rexp(200, rate = 7.3)
mean_7.3 = mean(exp.draws.7.3)
sd_7.3 = sd(exp.draws.7.3)

#lambda = 10
exp.draws.10 = rexp(200, rate = 10)
mean_10 = mean(exp.draws.10)
sd_10 = sd(exp.draws.10)

#Histogram of standard exponential distribution
hist(exp.draws.1)
hist_1 = (exp.draws.1)


```
   
For the histogram of the standard exponential distribution we can see that the the majority of the data is clustered between 0 and 1. This makes sense if we look at the summary of the data. We can see in the summary of the data that the median is median. This means that 50% of the data is below `r median_1` and 50% of the data is greater than `r median_1`. Therefore, in the histogram, it is sensible that ~130 out of 200 (~65%) randomly generated values would be in the bins between 0 and 1.    

```{r, echo=FALSE}
summary(exp.draws.1)
```

```{r, echo=FALSE}
#Plot values generated by the standard exponential dist. in the order they were generated
plot(exp.draws.1)

#Scatter plot between the standard distribution and the exponential distribution with lambda = 10
plot(exp.draws.1, exp.draws.10)

#Scatterplots from mean vector of previous distribution
mean_vector = c(mean_1, mean_0.2, mean_5, mean_7.3, mean_10)
sd_vector = c(sd_1, sd_0.2, sd_5, sd_7.3, sd_10)
rate_vector = c(1, 0.2, 5, 7.3, 10)
plot(rate_vector, mean_vector) #part a
plot(rate_vector, sd_vector) #part b
plot(sd_vector, mean_vector) #part c

#Part A:
big.exp.draws.1 = rexp(1100000)
big_mean = mean(big.exp.draws.1)
big_sd = sd(big.exp.draws.1)

#Part B:
hist(big.exp.draws.1)
x = seq(0, 15, by = 0.1)
plot(x, 1 - exp(-x))

#Part C:
mean(big.exp.draws.1[which(big.exp.draws.1 > 1)])

#Part D:
big.exp.draws.1.mat = matrix(data = big.exp.draws.1, nrow = 1100, ncol = 1000)
big_hist = hist(big.exp.draws.1.mat)

#Part E: 
col_371_mean = mean(big.exp.draws.1.mat[, 371])

#Part F:
col_means = colMeans(big.exp.draws.1.mat)
hist(col_means)

```


