---
title: "Nicole Pre-Class Week 08"
author: "Nicole Hardy"
date: "4 de noviembre de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(httr)
```

## Question 1: 

Read the HTML content of the following URL with a variable called webpage: https://money.cnn.com/data/us_markets/ At this point, it will also be useful to open this web page in your browser.

```{r}
#Question 1: reads html content
url = "https://money.cnn.com/data/us_markets/"
webpage = read_html(url)
print(webpage)
```

## Question 2: 

Get the session details (status, type, size) of the above mentioned URL.

```{r}
#Question 2: gets status, type, size of url
session_details = GET(url)
print(session_details)
```

## Question 3: 

Extract all of the sector names from the “Stock Sectors” table (bottom left of the web page.)

```{r}
#Question 3: extract all sector names of "Stock Sectors"
#xpath from selector gadget
webpage %>%
html_nodes(xpath = '//*[(@id = "wsod_sectorPerformance")]//*[contains(concat( " ", @class, " " ), concat( " ", "wsod_firstCol", " " ))]') %>%
html_text
```

## Question 4: 
Extract all of the “3 Month % Change” values from the “Stock Sectors” table.

```{r}
#Question 4: "Extract 3 Month % Change" values
#xpath from selector gadget
webpage %>%
  html_nodes(xpath = '//*[(@id = "wsod_sectorPerformance")]//td[contains(concat( " ", @class, " " ), concat( " ", "wsod_aRight", " " ))]') %>%
  html_text()
```

## Question 5: 
Extract the table “What’s Moving” (top middle of the web page) into a data-frame.
```{r}
#Question 5: Extracts table "What's Moving"
#xpath from inspect
webpage %>%
html_nodes(xpath = '//*[@id="wsod_marketMoversContainer"]/table') %>%
html_table()
```

## Question 6: 
Re-construct all of the links from the first column of the “What’s Moving” table. Hint: the base URL is “https://money.cnn.com”

```{r}
#Question 6: Builds links from first col of "What's Moving" Table
#xpath from selector gadget
base_url = "https://money.cnn.com"
stock_url = 
webpage %>%
html_nodes(xpath = '//*[(@id = "wsod_marketMoversContainer")]//*[contains(concat( " ", @class, " " ), concat( " ", "wsod_symbol", " " ))]') %>%
html_attr("href")
                    
#Reconstructs the links
paste(base_url, stock_url, sep = "")  
```

## Question 7: 
Extract the titles under the “Latest News” section (bottom middle of the web page.)

```{r}
#Question 7: Extracts titles under "Latest News"
#xpath from selector gadget
webpage %>%
  html_nodes(xpath = '//*[contains(concat( " ", @class, " " ), concat( " ", "HeadlineList", " " ))]//a') %>%
  html_text()    
```

## Question 8: 
To understand the structure of the data in a web page, it is often useful to know what the underlying attributes are of the text you see. Extract the attributes (and their values) of the HTML element that holds the timestamp underneath the “What’s Moving” table.

```{r}
#Question 8: Extract attributes of the timestamp underneath the "what's Moving Table"
#xpath from inspect
webpage %>%
  html_nodes(xpath = '//*[@id="wsod_marketMoversContainer"]/div/div/span') %>% 
  html_attrs()
```

## Question 9: 
Extract the values of the blue percentage-bars from the “Trending Tickers” table (bottom right of the web page.)
Hint: in this case, the values are stored under the “class” attribute.

```{r}
# Question 9: Extract percentage values of "Trending Tickers"  
webpage %>% 
  html_nodes(xpath = '//*[contains(concat( " ", @class, " " ), concat( " ", "bars", " " ))]') %>%
  html_attr(name = "class")
```

## Question 10: 
Get the links of all of the “svg” images on the web page.
```{r}
#Question 10: Get links of svg images
svg_url = 
webpage %>%
  html_nodes(xpath = '//*[contains(concat( " ", @class, " " ), concat( " ", "logo-cnn", " " ))]') %>%
  html_attr(name = "src")

#Builds links
paste(base_url, svg_url, sep = "")
```
