---
title: "Week 07"
author: "Nicole Hardy"
date: "23 de octubre de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library('rlist')
```

## Gambler's Ruin:
Gambler’s Ruin: Suppose you have a bankroll of $1000 and make bets of $100 on a fair game. By simulating the outcome directly for at most 5000 iterations of the game (or hands), estimate:

### Part A/D: 
The probability that you have “busted” (lost all your money) by the time you have placed your one hundredth bet. The mean and variance of your bankroll after 100 hands (including busts).

```{r, cache = TRUE}
#--------------------------------------------------------------
#Question 1:
#--------------------------------------------------------------
# Keeps track of wins and losses up to max number of bets or no bankroll
game_play = function(game, bankroll, num_bets, max_bets, bet)
{
  wins = 0
  losses = 0
  while (bankroll > 0 & num_bets < max_bets) 
  {
    roll = rbinom(1, 1, 0.5)
    if(roll == 1)
    {
      bankroll = bankroll + bet
      wins = wins + 1
      num_bets = num_bets + 1
    } else
    {
      bankroll = bankroll - bet
      losses = losses + 1
      num_bets = num_bets + 1
    }
  }
  game_matrix(game, wins, losses, bankroll)
}

# Saves a matrix with the individual trials as rows. 
#Saves in GLOBAL environment
game_matrix = function(entry, num_win, num_loss, money_left)
{
  trial[entry,1] <<- num_win + num_loss
  trial[entry,2] <<- num_win/trial[entry,1]
  trial[entry,3] <<- num_loss/trial[entry, 1]
  trial[entry, 4] <<- money_left
}

#Repeats everything for a set # of iterations. Saves results in matrix
#Prints probabilities, mean and variance
many_trials = function(bankroll, bet, num_bets, max_bets, iter, mean_fun, var_fun)
{
  for (game in 1:iter) 
  {
    bankroll2 = bankroll
    losses = 0
    wins = 0 
    game_play(game, bankroll2, num_bets, max_bets, bet)
  }
  return(c(calc_prob_busted(trial, iteration = iter), mean_time_busted(trial), 
           mean = mean_fun(trial$bankroll), var = var_fun(trial$bankroll)))
}

#Calculates probability of busting for a given number of maximum bets
calc_prob_busted = function(data, iteration)
{
  data %>%
    filter(bankroll <= 0) %>%
    summarise(busted_prob = n()/iteration)
}

#calculates the mean time you go bust
mean_time_busted = function(data)
{
  data %>%
    filter(bankroll <= 0) %>%
    summarise(mean_time = mean(total))
}

#Creates empty trial matrix saves to GLOBAL environment
empty_trial_mat = function(iter)
{
  trial <<- as.data.frame(matrix(data = 0, nrow = iter, ncol = 4))
  colnames(trial) <<- c("total", "prob of winning", "prob of losing", "bankroll")
}

#Creates empty matrix and runs everything for 5000 trials, 100 hands
empty_trial_mat(5000)
many_trials(1000, 100, 0, 100, 5000, mean, var)
head(trial)
```

###Part B/E: 
The probability that you have busted by the time you have placed your five hundredth bet by simulating the outcome directly. The mean and variance of your bankroll after 500 hands (including busts).

```{r}
#Creates empty matrix and runs everything for 5000 trials, 500 hands
empty_trial_mat(5000)
many_trials(1000, 100, 0, 500, 5000, mean, var)
head(trial)
```

### Part C: 
The mean time you go bust, given that you go bust within the first 5000 hands.

```{r}
#Mean time of bust 5000 hands
#Creates empty matrix and runs everything for 5000 trials, 5000 hands
empty_trial_mat(5000)
many_trials(1000, 100, 0, 5000, 5000, mean, var)
head(trial)
```

## Markov chains:
Suppose you have a game where the probability of winning on your first hand is 48%; each time you win, that probability goes up by one percentage point for the next game (to a maximum of 100%, where it must stay), and each time you lose, it goes back down to 48%. Assume you cannot go bust and that the size of your wager is a constant $100.

### Part a:
Is this a fair game? Simulate one hundred thousand sequential hands to determine the size of your return. Then repeat this simulation 99 more times to get a range of values to calculate the expectation.
```{r, cache = TRUE}
markov_play = function(game, bankroll, win_prob, num_bets, max_bets, bet)
{
  wins = 0
  losses = 0
  while (num_bets < max_bets & win_prob <= 1) 
  {
    roll = rbinom(1, 1, win_prob)
    if(roll == 1 & win_prob < 1)
    {
      bankroll = bankroll + bet
      wins = wins + 1
      num_bets = num_bets + 1
      win_prob = win_prob + 0.01
    } else if(roll == 0)
    {
      bankroll = bankroll - bet
      losses = losses + 1
      num_bets = num_bets + 1
      win_prob = 0.48
    } else if(roll == 1 & win_prob == 1)
    {
      bankroll = bankroll + bet
      wins = wins + 1
      num_bets = num_bets + 1
    }
  }
 game_matrix(game, wins, losses, bankroll)
}

markov_trials = function(bankroll, win_prob, bet, num_bets, max_bets, iter, mean_fun, var_fun)
{
  for (game in 1:iter) 
  {
    bankroll2 = bankroll
    losses = 0
    wins = 0 
    markov_play(game, bankroll2, win_prob, num_bets, max_bets, bet)
  }
  return(c(calc_prob_busted(trial, iteration = iter), mean_time_busted(trial), 
           mean = mean_fun(trial$bankroll), var = var_fun(trial$bankroll)))
}

empty_trial_mat(100)
markov_trials(1000, 0.48, 100, 0, 100000, 100, mean, var)
head(trial)
range(trial$bankroll)
```

No, this does not seem to be fair since the expectation is non-zero. 

### Part B:
Repeat this process but change the starting probability to a new value within 2% either way. Get the expected return after 100 repetitions. Keep exploring until you have a return value that is as fair as you can make it. Can you do this automatically?
```{r}
prob = seq(from = 0, to = 1, by = 0.1)
results = list()
for(i in 1:length(prob))
  {
    results[[i]] = markov_trials(1000, prob[i], 100, 0, 100000, 100, mean, var)
  }
```

The return value closest to a fair game is 82%. Yes, it can be done automatically. 

### Part C: 
Repeat again, keeping the initial probability at 48%, but this time change the probability increment to a value different from 1%. Get the expected return after 100 repetitions. Keep changing this value until you have a return value that is as fair as you can make it. 


