---
title: "Pre-Class Week 06"
author: "Nicole Hardy"
date: "14 de octubre de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1: 
1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

The first function check to see if the string begins with a certain prefix. I named it "starts_with_prefix."

The second function removes the last entry in the vector. I named it "removes_last."

The third function recycles vector y until a vector is outputted of length(x). I named it "recycle_y."

```
starts_with_prefix <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
remove_last <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
recycle_y <- function(x, y) {
  rep(y, length.out = length(x))
}
```

## Question 2: 
2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?

```{}
?rnorm
?MASS::mvrnorm
```

From the help we can see that rnorm() takes the arguments: n = # of observations, mean = vector of means, sd = vector of standard deviations. 

MASS:mvrnorm() takes the arguments: n = # of samples required, mu = a vector of means of the variables, Sigma = covariance matrix, tol = tolerance for lack of positive definitness in Sigma, empirical = logical where if true it Sigma specifies the empirical not population mean and covariance matrix, eispack = logical where values other than false are an error. 

The have in common n. However, their names for mean and standard deviations differ.

To be consistent I would assign the arguments the same names. For example instead of mu I would say mean. The name for standard deviation/Sigma differs because the multivariate case deals with a matrix of covariances (Sigma) and the univariate case deals with a vector of standard deviations (sd). They have different names because they are different data types, which would make it challenging to give these two arguments the same name. 

#Question 3:
3. Use lapply() and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset.

```{r}
lapply(mtcars, function(x) sd(x)/mean(x))
```

## Question 4:
### Part a:
4a. Use vapply() to compute the standard deviation of every column in a numeric data frame. 

```{r}
vapply(mtcars, sd, numeric(1))
```

### Part b:
4b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)

```{r}
vapply(iris[,vapply(iris, is.numeric, logical(1))], sd, numeric(1))
```
