---
title: "Pre-Class Week 05"
author: "Nicole Hardy"
date: "7 de octubre de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
suppressMessages(library(foreign))
```

## Question 1: Standardize Function
### Part A: 
A. Create a function called standardize.me() that takes a numeric vector as an argument, and returns the standardized version of the vector.

```{r}
#---------------------------------------------------------------------------
# Pre-Class Week 05: 
#---------------------------------------------------------------------------
#Loads data on coronary heart disease
wcgs <- read.dta("https://drive.google.com/uc?export=download&id=0B8CsRLdwqzbzYWxfN3ExQllBQkU")

#A function that standardizes the data to make each variable have
#mean = 0 and sd = 1
standardized.me = function(x)
{
  x = (x - mean(x, na.rm = TRUE))/sd(x, na.rm = TRUE)
  return(x)
}
```

### Part B: 
B. Assign all the numeric columns of the original WCGS dataset to a new dataset called WCGS.new.

```{r}
#Create new dataset with just numeric columns
WCGS.new = wcgs[,which(lapply(wcgs, class) == "numeric" | lapply(wcgs, class) == "integer")]
```

### Part C/D:
C. Using a loop and your new function, standardize all the variables WCGS.new dataset. 
D. What should the mean and standard deviation of all your new standardized variables be? Test your prediction by running a loop. 

The mean of the standardized data should be 0 and the standard deviation for the standardized variables should be 1. 

```{r}
#loops through dataset (minus id column) to standardize each column
for (i in 2:dim(WCGS.new)[2])
{
  WCGS.new[,i] = standardized.me(WCGS.new[,i])
  print(paste("Mean of column", i, "is", mean(WCGS.new[,i], na.rm = TRUE)))
  print(paste("Standard deviation of column", i, "is", sd(WCGS.new[,i], na.rm = TRUE)))
  print("") #to make it more readable. 
}
```

## Question 2: 
### Part A: 
A. Using a loop, calculate the mean weight of the subjects separated by the type of CHD they have.

```{r}
#calculate mean weight of subjects by type of CHD using a loop
types_CHD = as.character(unique(wcgs$typchd69))
for (i in types_CHD) 
{
  print(paste("Mean weight of person with",i, "is", mean(wcgs$weight[which(wcgs$typchd69 == i)], na.rm = TRUE)))
}
```

### Part B: 
B.  Now do the same thing, but now don’t use a loop

```{r}
wcgs %>%
  group_by(typchd69) %>%
  summarise(mean_weight = mean(weight))
```

